Measuring causal influence with\\ back-to-back regression

### 2019-05-23 paper/meg_exps.tex
meg results

The three types of analyses were implemented for each subject and each time
sample independently.  To estimate the reliability of these estimates in the
population, we tested, using one-sample $t$-test whether the distribution of
the estimated contribution of each factor different from baseline ($t=0$ms),
as $\text{t-test}(\hat E_{t=0} - \hat E_{\tau})$, with $\tau \in $ 150ms and
400ms respectively. We use the resulting p-values as proxy to quantify the
performance of each model.
% =======
% \paragraph{Back-to-back regression} We adapted back-to-back regression with two
% changes. First, to avoid getting a slightly positively-biased estimate of the
% factors contributing to the MEG signal, we swap the Tikhonov regularization of
% the first regression $G$ with a truncated regularization. Truncation was fitted
%  by maximum likelihood estimate, as implemented in scikit-learn principal
%  component analysis. Second, to minimize the variance increase induced by
%  bagging, we computed the inverse regularized covariance matrices outside
%  bagging:
% \begin{equation}
%   G = (Y^T Y+\lambda_Y I)^{-1} Y_{train}^T X_{train}
% \end{equation}
% \begin{equation}
% \hat X_{test} = Y_{test} \cdot G^T
% \end{equation} where $\left(train,test\right)$ are partitions of the data.
% %
% % \subsubsection{Second-level statistics}
% %
% % The three types of analyses were implemented for each subject and each time
% % sample independently.  To estimate the reliability of these estimates in the
% % population, we tested, using one-sample $t$-test whether the distribution of
% % the estimated contribution of each factor different from baseline ($t=0$ms),
% % as $\text{t-test}(\hat E_{t=0} - \hat E_{\tau})$, with $\tau \in $ 150ms and
% % 400ms respectively
% >>>>>>> 1775190a9737244f84b8ace529198a443029ad8d

\subsubsection{Results}
It is a well established fact that the number of letters in a string flashed
onto the retina modulates brain activity from 150ms after stimulus onset
\cite{pegado2014timing}. Similarly, word frequency is known to modulate brain
activity around 400ms \cite{kutas2011thirty}. Consequently, we aimed to compare
how sensitive each modeling approach in detecting these two types of brain
responses in a large neuroimaging dataset.

In Figure \ref{fig:meg_twocurves} we show for each model the effects obtained in a
representative subject, as well as the average effects obtained across subjects.
The results show that all methods significantly detect that both word length and
word frequency modulate brain responses after the stimulus onset. However
forward regression detected word length effects (XXX) and word frequency (XXX)
effects less reliably than B2BR (length=XXX, frequency=XXX). B2BR appeared to
be less sensitive than the backward regression (mean XXX). However, it is
important to highlight that backward models may mistakenly confound the two
factors and thus double the signal-to-noise ratio. This possibility is
corroborated by the remarkably similar dynamics between the two curves.

Finally we show in Figure \ref{fig:megresult} that B2BR is sufficiently sensitive
to reliably detect a multiplicity of subtle features, such as individual letters
and part-of-speech. This methods thus promises to improve the detectability of
subtle effects.

>>>>

The three types of analyses were implemented for each subject and each time
sample independently.  To estimate the reliability of these estimates in the
population, we tested, using one-sample $t$-test whether the distribution of
the estimated contribution of each factor different from baseline ($t=0$ms),
as $\text{t-test}(\hat E_{t=0} - \hat E_{\tau})$, with $\tau \in $ ranging from
150ms to 1s after word onset. We use the resulting p-values as proxy to quantify
the performance of each model. These results are depicted in Figure
\ref{fig:meg_pval}. For concision purposes, we only report this metric averaged
between 150ms and 1s after word onset.
% =======
% \paragraph{Back-to-back regression} We adapted back-to-back regression with two
% changes. First, to avoid getting a slightly positively-biased estimate of the
% factors contributing to the MEG signal, we swap the Tikhonov regularization of
% the first regression $G$ with a truncated regularization. Truncation was fitted
%  by maximum likelihood estimate, as implemented in scikit-learn principal
%  component analysis. Second, to minimize the variance increase induced by
%  bagging, we computed the inverse regularized covariance matrices outside
%  bagging:
% \begin{equation}
%   G = (Y^T Y+\lambda_Y I)^{-1} Y_{train}^T X_{train}
% \end{equation}
% \begin{equation}
% \hat X_{test} = Y_{test} \cdot G^T
% \end{equation} where $\left(train,test\right)$ are partitions of the data.
% %
% % \subsubsection{Second-level statistics}
% %
% % The three types of analyses were implemented for each subject and each time
% % sample independently.  To estimate the reliability of these estimates in the
% % population, we tested, using one-sample $t$-test whether the distribution of
% % the estimated contribution of each factor different from baseline ($t=0$ms),
% % as $\text{t-test}(\hat E_{t=0} - \hat E_{\tau})$, with $\tau \in $ 150ms and
% % 400ms respectively
% >>>>>>> 1775190a9737244f84b8ace529198a443029ad8d

\subsubsection{Results}
The number of letters in a string flashed onto the retina is known to modulate
brain activity from 150ms after stimulus onset \cite{pegado2014timing}.
Similarly, word frequency is known to modulate brain activity around 400ms \cite{kutas2011thirty}. Consequently, we aimed to compare how sensitive each
model is in detecting these two types of brain responses in a large
neuroimaging dataset, where word length and word frequency were collinear.

In Figure \ref{fig:meg_twocurves} we show, for each model, the effects obtained
in a representative subject, as well as the average effects across subjects.
The results show that all models significantly detect that both word length and
word frequency modulate brain responses after the stimulus onset. However
forward regression detected word length ($t-value=10.6$, $p=10^{-16}$) and
word frequency ($t-value=11.1$, $p=10^-17$) less reliably than B2B (length:
$t-value=18.7$, $p=10^{-31}, frequency:$t-value=15.9$, $p=10^{-26}$). B2B
was less sensitive than the backward regression (length: $t-value=33.6$,
$p=10^{-50}$, $t-value=31.7$, frequency:$p=10^{-48}$). However, as detailed in
the introduction, backward models confound collinear factors. Consequently,
the high sensitivity of our backward regression is likely to come from a loss of specificity. This possibility is corroborated by the fact that the dynamics
associated with word length and word frequency are remarkably similar in the
backward model.

Finally we investigate in Figure \ref{fig:megresult} whether B2B was sufficiently sensitive to reliably detect a multiplicity of subtle features, such as individual letters and part-of-speech. All of these features appeared to be reliably detected \cite{fig:megbar}. B2B thus promises to improve the detectability of subtle effects.
